---
title: "Complete intersectr demo"
author: "dblodgett@usgs.gov"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>",
fig.width=6,
fig.height=4
)
options(scipen = 9999)
library(magrittr)
library(intersectr)
```

This is a complete demo showing a workflow implemented with [`intersectr`](https://github.com/dblodgett-usgs/intersectr). One of the primary goals of the package is to intersect gridded time series data with polygon geometries. For example, and as shown below, we can get time-series of area-weighed average monthly precipitation for counties in North Carolina from the PRISM source dataset that has a 4km - monthly resolution.

The example here uses a dataset from a THREDDS/OPeNDAP data server. Local NetCDF files can also be used but may require special handling if files are broken up through time.

The [`ncmeta`](https://github.com/hypertidy/ncmeta/) package provides most of the NetCDF metadata handling and will work for either remote OPeNDAP hosted data or local NetCDF data. Below, we get a list of variables then find the coordinate variables for a selected variable.
```{r nc_setup}
nc_file <- "https://cida.usgs.gov/thredds/dodsC/prism_v2"
(nc_var <- ncmeta::nc_vars(nc_file))
variable_name <- "ppt"
(nc_coord_vars <- ncmeta::nc_coord_var(nc_file, variable_name))
```

We'll use the sample polygon provided by the [`sf`](https://github.com/r-spatial/sf) package as our geometry. The EPSG:5070 projection is a reasonable equal area projection for work in the Conterminous US.
```{r geom}
(geom <- sf::st_transform(sf::read_sf(system.file("shape/nc.shp",
                                                  package = "sf")),
                          "+init=epsg:5070"))
plot(sf::st_geometry(geom))
```

  Now we can create cell geometry for the grid we are going to intersect with our county geometry. A function from [`ncmeta`](https://github.com/hypertidy/ncmeta/blob/master/R/nc-prj.R) helps figure out the right proj4 string from our NetCDF file.

While very small, some of the row coordinates spacings are big enough to be interpreted as irregular in the create_cell_geometry function. We can even them out with a regular sequence. 

The create_cell_geometry function generates an `sf` `data.frame` representing each cell from the dataset as a polygon with a unique id.
```{r cells}
(nc_prj <- ncmeta::nc_gm_to_prj(ncmeta::nc_grid_mapping_atts(nc_file)))

nc <- RNetCDF::open.nc(nc_file)
X_coords <- RNetCDF::var.get.nc(nc, nc_coord_vars$X, unpack = TRUE)
Y_coords <- RNetCDF::var.get.nc(nc, nc_coord_vars$Y, unpack = TRUE)

(cell_geometry <-
    create_cell_geometry(X_coords = X_coords,
                         Y_coords = Y_coords,
                         prj = nc_prj,
                         geom = geom, 
                         buffer_dist = 10000,
                         regularize = TRUE))

plot(sf::st_geometry(sf::st_transform(cell_geometry, 
                                      sf::st_crs(geom))), lwd = 0.2)
plot(sf::st_geometry(geom), add = TRUE)
```

With our input geometries ready, we can run the area_weights and execute_intersection functions. 

The area weights function provides weights for area-weighted statistics. The weight column of the output can be used in an [area-weighted-mean](https://en.wikipedia.org/wiki/Weighted_arithmetic_mean) calculation where mean of the data source cells is desired for each of the county polygons.

```{r area_weights}
data_source_cells <- sf::st_sf(dplyr::select(cell_geometry, grid_ids))
target_polygons <- sf::st_sf(dplyr::select(geom, FIPS))
sf::st_agr(data_source_cells) <- "constant"
sf::st_agr(target_polygons) <- "constant"
area_weights <- calculate_area_intersection_weights(
  data_source_cells,
  target_polygons)
```

Now we have everything we need to actually run the intersection. This function iterates through time steps, applying the intersection weights for each time step.
```{r run_intersection}
start_date <- "2010-01-01 00:00:00"
end_date <- "2011-01-01 00:00:00"
intersected <- execute_intersection(nc_file = nc_file,
                                    variable_name = variable_name,
                                    intersection_weights = area_weights,
                                    cell_geometry = cell_geometry, 
                                    x_var = nc_coord_vars$X,
                                    y_var = nc_coord_vars$Y,
                                    t_var = nc_coord_vars$T, 
                                    start_datetime = start_date, 
                                    end_datetime = end_date)
# Need to fix in execute_intersection
intersected[, 2:ncol(intersected)] <- intersected[, 2:ncol(intersected)]
intersected_intersectr <- intersected
```

Now we can plot up some of the resulting data on a map or through time.
```{r plot_result}
row <- which(as.character(intersected$time_stamp) == "2010-06-01")
geom_data <- dplyr::select(geom, FIPS) %>%
  dplyr::left_join(data.frame(FIPS = names(intersected)[2:ncol(intersected)],
                              poly_data = as.numeric(intersected[row, 2:ncol(intersected)]),
                              stringsAsFactors = FALSE),
                   by = "FIPS")
plot(geom_data["poly_data"],
     main = paste(intersected[row, 1], "Prism Precip mm/month"))
plot(intersected$time_stamp, seq(min(intersected[, 2:ncol(intersected)]), 
                                 max(intersected[, 2:ncol(intersected)]),
                                 length.out = nrow(intersected)), col = NA, 
     xlab = "Year", ylab = "PRISM Indicated Precip (mm/month)")
lines(intersected$time_stamp, apply(intersected[, 2:ncol(intersected)], 1, mean), lwd = 2, col = "grey")
lines(intersected$time_stamp, apply(intersected[, 2:ncol(intersected)], 1, max), lwd = .5, col = "grey")
lines(intersected$time_stamp, apply(intersected[, 2:ncol(intersected)], 1, min), lwd = .5, col = "grey")
```

```{r geoknife}
library(geoknife)
fabric <- webdata('prism')
times(fabric) <- c(start_date, end_date)
variables(fabric) <- "ppt"
stencil <- webgeom(geom = "sample:Counties", attribute = "FIPS")
values(stencil) <- geom$FIPS
job <- geoknife(stencil, fabric, wait = TRUE)
intersected <- result(job)
intersected <- intersected[, 1:(ncol(intersected) - 2)]
names(intersected)[1] <- "time_stamp"
intersected_geoknife <- intersected
```

```{r plot_results_2} 
row <- which(as.character(intersected$time_stamp) == "2010-06-01")
geom_data <- dplyr::select(geom, FIPS) %>%
  dplyr::left_join(data.frame(FIPS = names(intersected)[2:ncol(intersected)],
                              poly_data = as.numeric(intersected[row, 2:ncol(intersected)]),
                              stringsAsFactors = FALSE),
                   by = "FIPS")
plot(geom_data["poly_data"],
     main = paste(intersected[row, 1], "Prism Precip mm/month"))
plot(intersected$time_stamp, seq(min(intersected[, 2:ncol(intersected)]), 
                                 max(intersected[, 2:ncol(intersected)]),
                                 length.out = nrow(intersected)), col = NA, 
     xlab = "Year", ylab = "PRISM Indicated Precip (mm/month)")
lines(intersected$time_stamp, apply(intersected[, 2:ncol(intersected)], 1, mean), lwd = 2, col = "grey")
lines(intersected$time_stamp, apply(intersected[, 2:ncol(intersected)], 1, max), lwd = .5, col = "grey")
lines(intersected$time_stamp, apply(intersected[, 2:ncol(intersected)], 1, min), lwd = .5, col = "grey")
```

```{r plot}
X_inds <- seq(min(cell_geometry$X_ind), max(cell_geometry$X_ind), 1)
Y_inds <- seq(min(cell_geometry$Y_ind), max(cell_geometry$Y_ind), 1)
ids <- intersectr:::get_ids(length(X_inds), length(Y_inds))
dates <- RNetCDF::var.get.nc(nc, "time", unpack = TRUE)
date_units <- RNetCDF::att.get.nc(nc, "time", "units")
dates <- RNetCDF::utcal.nc(date_units, dates, type = "c")
date_ind <- which(as.character(dates) == "2010-06-01")
grid_data <- RNetCDF::var.get.nc(nc, variable_name,
                                 start = c(min(X_inds), min(Y_inds), date_ind),
                                 count = c(length(X_inds), length(Y_inds), 1), 
                                 unpack = TRUE)
grid_data <- data.frame(grid_data = matrix(grid_data,
                                           ncol = 1,
                                           byrow = TRUE),
                        grid_ids = matrix(ids, ncol = 1))
grid_data$grid_data[grid_data$grid_data < 0] <- NA
grid_data <- dplyr::left_join(cell_geometry, grid_data, by = "grid_ids")
intersected <- intersected_intersectr
row <- which(as.character(intersected$time_stamp) == "2010-06-01")
geom_data <- dplyr::select(geom, FIPS) %>%
  dplyr::left_join(data.frame(FIPS = names(intersected)[2:ncol(intersected)],
                              poly_data = as.numeric(intersected[row, 2:ncol(intersected)]), 
                              stringsAsFactors = FALSE),
                   by = "FIPS")
geom_data <- sf::st_transform(geom_data, sf::st_crs(grid_data))
breaks <- c(0, 80, 120, 200, 300, 400)
plot(grid_data["grid_data"], border = NA, breaks = breaks)
plot(geom_data["poly_data"], breaks = breaks)
plot(grid_data$geometry)
plot(grid_data["grid_data"], border = NA, breaks = breaks, add = TRUE)
plot(geom_data["poly_data"], breaks = breaks, add = TRUE)
```

```{r diff}
row_gk <- which(as.character(intersected_geoknife$time_stamp) == "2010-06-01")
row_ir <- which(as.character(intersected_intersectr$time_stamp) == "2010-06-01")
diffs <- (intersected_intersectr[row_ir, 2:ncol(intersected_intersectr)] - 
            intersected_geoknife[row_gk, 2:ncol(intersected_geoknife)]) / 
  ((intersected_intersectr[row_ir, 2:ncol(intersected_intersectr)] + 
      intersected_geoknife[row_gk, 2:ncol(intersected_geoknife)]) / 2)
max(as.numeric(diffs[1,]))
mean(as.numeric(diffs[1,]))
```
