---
title: "Comparison with geoknife and the Geo Data Portal"
author: "dblodgett@usgs.gov"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
  ```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
  )
  options(scipen = 9999)
  library(magrittr)
  library(intersectr)
  ```
  
  This is a complete demo showing a workflow implemented with [`intersectr`](https://github.com/dblodgett-usgs/intersectr) compared to the same workflow implemented with [`geoknife`](http://usgs-r.github.io/geoknife/). This both demonstrates `intersectr` and shows that we get the same answers as the well-vetted [Geo Data Portal](https://cida.usgs.gov/gdp/), as used by `geoknife`.
  
  One of the primary goals of the package is to intersect gridded time series data with polygon geometries. For example, and as shown below, we can get time-series of area-weighed average monthly precipitation for counties in North Carolina from the PRISM source dataset that has a 4km - monthly resolution.
  
  The example here uses a dataset from a THREDDS/OPeNDAP data server. Local NetCDF files can also be used but may require special handling if files are broken up through time.
  
  The [`ncmeta`](https://github.com/hypertidy/ncmeta/) package provides most of the NetCDF metadata handling and will work for either remote OPeNDAP hosted data or local NetCDF data. Below, we get a list of variables and their coordinate variables that will be used later to construct cell geometry.
  ```{r nc_setup}
  library(ncmeta)
  library(dplyr)
  library(RNetCDF)
  library(sf)
  nc_file <- "https://cida.usgs.gov/thredds/dodsC/prism_v2"
  (nc_coord_vars <- nc_coord_var(nc_file))
  variable_name <- "ppt"
  nc_coord_vars <- filter(nc_coord_vars, variable == variable_name)
  ```
  
  We'll use the same sample polygons provided by the [`sf`](https://github.com/r-spatial/sf) package as our geometry. To make sure we get a good comparison with the Geo Data Portal, a more detailed version of the North Carolina county geometry is available as part of `intersectr`.
  ```{r geom}
geom <- read_sf(system.file("extdata/detailed_counties.gpkg", 
package = "intersectr"))
plot(st_geometry(geom))
```

Now we can create cell geometry for the grid we are going to intersect with our county geometry. A function from [`ncmeta`](https://github.com/hypertidy/ncmeta/blob/master/R/nc-prj.R) helps figure out the right proj4 string from our NetCDF file.

While very small, some of the row coordinates spacings are big enough to be interpreted as irregular in the create_cell_geometry function. We can even them out with the `regularize = TRUE` input. 

The create_cell_geometry function generates an `sf` `data.frame` representing each cell from the dataset as a polygon with a unique id.
```{r cells}
(nc_prj <- nc_gm_to_prj(nc_grid_mapping_atts(nc_file)))

nc <- open.nc(nc_file)
X_coords <- var.get.nc(nc, nc_coord_vars$X, unpack = TRUE)
Y_coords <- var.get.nc(nc, nc_coord_vars$Y, unpack = TRUE)

(cell_geometry <-
    create_cell_geometry(X_coords = X_coords,
                         Y_coords = Y_coords,
                         prj = nc_prj,
                         geom = geom, 
                         buffer_dist = 0.1, # Degrees
                         regularize = TRUE))

plot(st_geometry(st_transform(cell_geometry, 
                              st_crs(geom))), lwd = 0.2)
plot(st_geometry(geom), add = TRUE)
```

The area weights function provides weights for area-weighted statistics. The weight column of the output can be used in an [area-weighted-mean](https://en.wikipedia.org/wiki/Weighted_arithmetic_mean) calculation where mean of the data source cells is desired for each of the county polygons.

```{r area_weights}
data_source_cells <- st_sf(select(cell_geometry, grid_ids))
target_polygons <- st_sf(select(geom, FIPS))
st_agr(data_source_cells) <- "constant"
st_agr(target_polygons) <- "constant"
area_weights <- calculate_area_intersection_weights(
  data_source_cells,
  target_polygons, allow_lonlat = TRUE)
```

Now we have everything we need to actually run the intersection. The `execute_intersection` function iterates through time steps, applying the intersection weights for each time step.
```{r run_intersection}
start_date <- "2010-01-01 00:00:00"
end_date <- "2011-01-01 00:00:00"
intersected <- execute_intersection(nc_file = nc_file,
                                    variable_name = variable_name,
                                    intersection_weights = area_weights,
                                    cell_geometry = cell_geometry, 
                                    x_var = nc_coord_vars$X,
                                    y_var = nc_coord_vars$Y,
                                    t_var = nc_coord_vars$T, 
                                    start_datetime = start_date, 
                                    end_datetime = end_date)

# Save for later comparison.
intersected_intersectr <- intersected
```

Now we can plot up some of the resulting data on a map or through time.
```{r plot_result}
date <- "2010-06-01"
row <- which(as.character(intersected_intersectr$time_stamp) == date)
ncols <- ncol(intersected_intersectr)
fips <- names(intersected)[2:ncols]
ts_data <- intersected_intersectr[, 2:ncols]
poly_data <- as.numeric(intersected_intersectr[row, 2:ncols])
ts_range <- seq(min(ts_data), max(ts_data), length.out = nrow(ts_data))
time_stamps <- intersected_intersectr$time_stamps

geom_data <- geom %>%
  select(FIPS) %>%
  left_join(data.frame(FIPS = fips, poly_data = poly_data,
                       stringsAsFactors = FALSE), by = "FIPS")

plot(geom_data["poly_data"], main = paste(date, "Prism Precip mm/month"))
plot(time_stamps, ts_range, col = NA, 
     xlab = "Year", ylab = "PRISM Indicated Precip (mm/month)")
lines(time_stamps, apply(ts_data, 1, mean), lwd = 2, col = "grey")
lines(time_stamps, apply(ts_data, 1, max), lwd = .5, col = "grey")
lines(time_stamps, apply(ts_data, 1, min), lwd = .5, col = "grey")
```

Now lets run the same thing with `geoknife` and plot up the same thing.
```{r geoknife, eval=FALSE}
library(geoknife)
fabric <- webdata('prism')
times(fabric) <- c(start_date, end_date)
variables(fabric) <- "ppt"

stencil <- webgeom(geom = "sample:Counties", attribute = "FIPS")
values(stencil) <- target_polygons$FIPS

job <- geoknife(stencil, fabric, wait = TRUE)

intersected <- result(job)
intersected <- intersected[, 1:(ncol(intersected) - 2)]
names(intersected)[1] <- "time_stamps"

order <- match(names(intersected_intersectr), names(intersected))

intersected_geoknife <- intersected[, order]
```
```{r load_cached_data, echo=FALSE}
intersected_geoknife <- readRDS(system.file("extdata/intersected_geoknife.rds",
                                            package = "intersectr"))
```
```{r plot_results_2} 
row <- which(as.character(intersected_geoknife$time_stamp) == date)
ncols <- ncol(intersected_geoknife)
fips <- names(intersected_geoknife)[2:ncols]
ts_data <- intersected_geoknife[, 2:ncols]
poly_data <- as.numeric(intersected_geoknife[row, 2:ncols])
ts_range <- seq(min(ts_data), max(ts_data), length.out = nrow(ts_data))
time_stamps <- intersected_geoknife$time_stamps

geom_data <- geom %>%
  select(FIPS) %>%
  left_join(data.frame(FIPS = fips, poly_data = poly_data,
                       stringsAsFactors = FALSE), by = "FIPS")

plot(geom_data["poly_data"], main = paste(date, "Prism Precip mm/month"))
plot(time_stamps, ts_range, col = NA, 
     xlab = "Year", ylab = "PRISM Indicated Precip (mm/month)")
lines(time_stamps, apply(ts_data, 1, mean), lwd = 2, col = "grey")
lines(time_stamps, apply(ts_data, 1, max), lwd = .5, col = "grey")
lines(time_stamps, apply(ts_data, 1, min), lwd = .5, col = "grey")

```

To get an idea of what this looks like on a map, now we can plot up the grid data and polygon data together.
```{r plot}

X_inds <- seq(min(cell_geometry$X_ind), max(cell_geometry$X_ind), 1)
Y_inds <- seq(min(cell_geometry$Y_ind), max(cell_geometry$Y_ind), 1)

ids <- intersectr:::get_ids(length(X_inds), length(Y_inds))

dates <- utcal.nc(att.get.nc(nc, "time", "units"), 
                  var.get.nc(nc, "time", unpack = TRUE), 
                  type = "c")

date_ind <- which(as.character(dates) == date)

grid_data <- var.get.nc(nc, variable_name,
                        start = c(min(X_inds), min(Y_inds), date_ind),
                        count = c(length(X_inds), length(Y_inds), 1), 
                        unpack = TRUE)
grid_data <- data.frame(grid_data = matrix(grid_data,
                                           ncol = 1,
                                           byrow = TRUE),
                        grid_ids = matrix(ids, ncol = 1))
grid_data$grid_data[grid_data$grid_data < 0] <- NA
grid_data <- left_join(cell_geometry, grid_data, by = "grid_ids")
intersected <- intersected_intersectr
row <- which(as.character(intersected$time_stamp) == "2010-06-01")
geom_data <- select(geom, FIPS) %>%
  left_join(data.frame(FIPS = names(intersected)[2:ncol(intersected)],
                       poly_data = as.numeric(intersected[row, 2:ncol(intersected)]), 
                       stringsAsFactors = FALSE),
            by = "FIPS")
geom_data <- st_transform(geom_data, st_crs(grid_data))
breaks <- c(0, 80, 120, 200, 300, 400)
plot(grid_data["grid_data"], border = NA, breaks = breaks)
plot(geom_data["poly_data"], breaks = breaks)
plot(grid_data$geometry)
plot(grid_data["grid_data"], border = NA, breaks = breaks, add = TRUE)
plot(geom_data["poly_data"], breaks = breaks, add = TRUE)
```  

Now lets look at how these two results compare. As the historgram and map shows below, all differences are below 1%!
```{r diff}
row_ir <- which(as.character(intersected_intersectr$time_stamp) == "2010-06-01")
row_gk <- which(as.character(intersected_geoknife$time_stamp) == "2010-06-01")
data_ir <- intersected_intersectr[row_ir, 2:ncol(intersected_intersectr)]
data_gk <- intersected_geoknife[row_gk, 2:ncol(intersected_geoknife)]
all(names(data_ir) == names(data_gk))
compare <- data.frame(fips = names(data_ir), 
                      ir = as.numeric(data_ir), 
                      gk = as.numeric(data_gk), stringsAsFactors = FALSE)
compare["diffs"] <- 100 * (as.numeric((data_ir - data_gk) / ((data_ir + data_gk) / 2)))

hist(compare$diffs, 
     main = "Histogram of percent difference intersectr - geoknife.", 
     xlab = "Percent Different")
```

```{r which_diff}
max_diff <- which(abs(compare$diffs) == max(abs(compare$diffs)))
data_ir[max_diff]
data_gk[max_diff]

geom_data <- left_join(select(geom_data, FIPS, poly_data),
                       select(compare, fips, diffs), by = c("FIPS" = "fips"))
plot(geom_data["diffs"])
```
